Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-09-06 doc/manuscript/manuscript.Rmd

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each genotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis. 
We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].

@@

### Analysis of Shared Multilocus Genotypes {-}

We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could be present in more than one population, that MLG would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version `r packageVersion("ggplot2")` [@ggplot2] and *ggraph* `r packageVersion("ggraph")` [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the five polymorphic markers described above.

@@

![
Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
Singletons (46) were truncated, leaving 41 MCGs.
White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
Details in fig. \@ref(fullgraph). 
**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
Filled nodes represent MLGs and open nodes represent MCGs.
Numbers inside nodes are the MLG/MCG label. 
Node (circle) area scaled to the number of samples represented (range: 1--73).
Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
Edges extending from MLGs displayed to other MCGs are not shown.
](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}

@@

Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49') (Table \@ref(tab:MCG-aggressiveness)).
We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140') (Table \@ref(tab:MLG-aggressiveness)).

### Correlation of Mulitlocus Genotypes and Mycelial Compatibility Groups {-}

Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLG unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig. \@ref(fullgraph)).
The three communities with the most members contained eight of the 10 most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG. 
Representing just four isolates, 'MLG 75' had `r big_genotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
Overall, there was no clear pattern to the association between MLG and MCGs.

## Structure of Shared Multilocus Genotypes {-}

We observed a total of 165 MLGs (215 with 16 loci).
The most abundant MLG was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
Within regions, the genotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLGs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.

@@

```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
read_csv("../../results/tables/mlg_table.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(.default = col_character())) %>%
  rename(`$P_{sex}$` = Psex) %>% 
  knitr::kable(format = "pandoc")
```

Table: (\#tab:mlg-table) The 5 most abundant Multilocus Genotypes (MLG) with
the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.

@@

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions. 
We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the wmn can be effective at screening for potential resistant lines within growing regions.

@@

Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (`r andlist(c(names(strata(dat11)), "Aggressiveness"))`) [@legendre1999distance].
From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
When we inspected the biplot of the initial results, we saw that the most important predictors were 'MCG 44', 'MCG 5', and 'MCG 9'. 
We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig. \@ref(biggraph)).
We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexico isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
Aggressiveness and specific source field had little to no effect on genetic diversity. 
These results are in agreement with studies that examined differentiation based on Host [@aldrich-wolfe2015genetic] and Aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables. 
This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.

## Mycelial Compatibility Groups and Multilocus Genotypes {-}

Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
Our results do not support this assumption.
It can be seen in Fig. \@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG ('MLG 25').
What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig. \@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_genotype$N` other genotypes through three MCG when the population structure of *S. sclerotiorum* is known to be clonal.

Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic]. 
Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLGs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].

@@

The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone. 
This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.

The widespread nature of multilocus genotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
It has previously been shown that infections can be transmitted through seed [@botelho2013performance], but seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management].
Thus, we hypothesize that shared genotypes between populations may arise due to transmission events of seed or sclerotia.
However, because we observe population structure based on Region, we speculate that these transmission events are rare.

@@

Another limitation was the microsatellite markers used for this particular study [@sirjusingh2001characterisation].
The genotype accumulation curve showed no indication of a plateau, indicating that if we had sampled more loci, we would have resolved more multilocus genotypes. 
While 16 loci showed us similar results and began to show a plateau for the genotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci. 
With the availability of an optically-mapped genome [@derbyshire2017complete], future studies describing the genetic diversity of *S. sclerotiorum* should employ techniques such as Genotyping-By-Sequencing [@davey2011genome], Sequence Capture [@grover2012targeted], or Whole Genome Sequencing.

@@

This study represents the largest genetic analysis of *S. sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
Based on our analysis of the relationships between MCG and MLG, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones. 
Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S. sclerotiorum* may be considered representative of the surrounding regions. 
While we found no direct relationship between genotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.

@@

```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
read_csv("../../results/tables/MLG-aggressiveness.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(.default = col_character())) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 40)
```

Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLG; groupings according to 95% family-wise confidence interval.

![
Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Genotypes (MLG) (full nodes) where the number in each node represents the MLG/MCG assignment. 
Node size reflect the number of samples represented by each node (circle). 
Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of samples shared. 
Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}

@@

![
Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 16 loci. 
Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
](../../results/figures/publication/mlg-16.pdf){#community-graph16 width=100%}

>>>>

All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
<!-- add something about repeat length consistency  -->
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each haplotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis. 
We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].

@@

### Analysis of Shared Multilocus Haplotypes {-}

We wanted to evaluate patterns of connectivity between shared multilocus haplotypes across geographic regions.
We first tabulated the multilocus haplotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared haplotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
Each node was weighted by the fraction of shared MLHs.
Each edge represented a single MLH, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same haplotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version `r packageVersion("ggplot2")` [@ggplot2] and *ggraph* `r packageVersion("ggraph")` [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the five polymorphic markers described above.

@@

![
Data show complex associations between Mycelial Compatibility Groups and Multilocus Haplotypes.
**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
Singletons (46) were truncated, leaving 41 MCGs.
White bars represent sample counts and grey bars represent counts of unique multilocus haplotypes (MLH).
The transparency of the bars represent the evenness of the distribution of the MLHs within a given MCG.
A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
**B)** Full graph-representation of the relationship between MCGs (open circles) and MLHs (filled circles).
Details in fig. \@ref(fullgraph). 
**C)** A subset of **B** representing the 8 most common MCGs and their associated MLHs (dashed box in **A**).
Filled nodes represent MLHs and open nodes represent MCGs.
Numbers inside nodes are the MLH/MCG label. 
Node (circle) area scaled to the number of samples represented (range: 1--73).
Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
Edges extending from MLHs displayed to other MCGs are not shown.
](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}

@@

Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49') (Table \@ref(tab:MCG-aggressiveness)).
We additionally found a significant effect for MLHs (*P* = 7.44e^-4^) for MLHs, with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).

### Correlation of Mulitlocus Haplotypes and Mycelial Compatibility Groups {-}

Our network-based approach to correlating MLHs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLH unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLHs are unable to cross with any other MLH in this data set (Fig. \@ref(fullgraph)).
The three communities with the most members contained eight of the 10 most abundant MCGs. 
Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH. 
Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
Overall, there was no clear pattern to the association between MLH and MCGs.

## Structure of Shared Multilocus Haplotypes {-}

We observed a total of 165 MLHs (215 with 16 loci).
The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
Within regions, the haplotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
Of the 165 MLHs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17 (Fig. \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig. \@ref(community-graph16)).
The international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.

@@

```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
read_csv("../../results/tables/mlg_table.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(.default = col_character())) %>%
  rename(`$P_{sex}$` = Psex) %>% 
  rename(MLH = MLG) %>%
  knitr::kable(format = "pandoc")
```

Table: (\#tab:mlg-table) The 5 most abundant Multilocus Haplotypes (MLH) with
the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.

@@

In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions. 
We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared haplotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico. 
To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
These findings indicate that the wmn can be effective at screening for potential resistant lines within growing regions.

@@

Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (`r andlist(c(names(strata(dat11)), "Aggressiveness"))`) [@legendre1999distance].
From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
When we inspected the biplot of the initial results, we saw that the most important predictors were 'MCG 44', 'MCG 5', and 'MCG 9'. 
We believe that this is driven by the fact that these particular MCGs have uneven MLH distributions, meaning that they are heavily associated with one particular MLH (Fig. \@ref(biggraph)).
We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexico isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
Aggressiveness and specific source field had little to no effect on genetic diversity. 
These results are in agreement with studies that examined differentiation based on Host [@aldrich-wolfe2015genetic] and Aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables. 
This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.

## Mycelial Compatibility Groups and Multilocus Haplotypes {-}

Of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
Our results do not support this assumption.
It can be seen in Fig. \@ref(biggraph)A that the most abundant MCG contains several MLHs, but the diversity of those MLHs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLH ('MLH 25').
What is not shown in Fig. \@ref(biggraph)A is the MLHs that are shared between MCGs. 
This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig. \@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S. sclerotiorum* is known to be clonal.

Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic]. 
Either several MCGs belong to one MLH, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].

@@

The results show several haplotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone. 
This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.

The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
It has previously been shown that infections can be transmitted through seed [@botelho2013performance], but seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management].
Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
However, because we observe population structure based on Region, we speculate that these transmission events are rare.

@@

Another limitation was the microsatellite markers used for this particular study [@sirjusingh2001characterisation].
The haplotype accumulation curve showed no indication of a plateau, indicating that if we had sampled more loci, we would have resolved more multilocus haplotypes. 
While 16 loci showed us similar results and began to show a plateau for the haplotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci. 
With the availability of an optically-mapped genome [@derbyshire2017complete], future studies describing the genetic diversity of *S. sclerotiorum* should employ techniques such as Genotyping-By-Sequencing [@davey2011genome], Sequence Capture [@grover2012targeted], or Whole Genome Sequencing.

@@

This study represents the largest genetic analysis of *S. sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones. 
Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S. sclerotiorum* may be considered representative of the surrounding regions. 
While we found no direct relationship between haplotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.

@@

```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
read_csv("../../results/tables/MLG-aggressiveness.csv", 
         na = "NA", # spaces read in as spaces
         col_types = cols(.default = col_character())) %>%
  rename(MLH = MLG) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 40)
```

Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLH; groupings according to 95% family-wise confidence interval.

![
Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes (MLH) (full nodes) where the number in each node represents the MLH/MCG assignment. 
Node size reflect the number of samples represented by each node (circle). 
Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of samples shared. 
Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}

@@

![
Network of populations (nodes/circles) and their shared multilocus haplotypes (MLH) (edges/lines) haplotyped over 16 loci. 
Each node is labeled with **name (number of MLHs shared/number of MLHs total).** 
The shade and area of the nodes are proportional to the number of unique MLHs within the node and the inner nodes are proportional to the number of private MLHs to the region (bottom legend). 
Each edge represents a single MLH where its thickness represents the number of populations that share the MLH and the shade represents the value of $P_{sex}$, or the probability of encountering that MLH from two independent meiotic events.
](../../results/figures/publication/mlg-16.pdf){#community-graph16 width=100%}
